---
title: "Models of Neighborhood Change in Seattle, WA"
output: 
  flexdashboard::flex_dashboard:
    theme: bootstrap
    source: embed
    smart: false
runtime: shiny
---

    

```{r global, echo=F}

# PACKAGES 


# dashboard layout, widgets, and uploading

library( flexdashboard )
library( shiny )
library( rsconnect )

# data wrangling 

library( dplyr )


# formatting output

library( DT )
library( pander )
library( knitr )
library( stargazer )


# maps

library( ggmap )
library( leaflet )
library( viridis )
library( geojsonio )
library( sp )
library( sf )
library( tmap )
library( pals )

```

```{r}
# DATA STEPS 

# load dorling cartogram from github
# map already contains census data and groups from clustering 

github.url <- "https://raw.githubusercontent.com/drewradovich/ASU-MSDS-Projects/main/sea_dorling.geojson"
sea <- geojson_read( x=github.url,  what="sp" )

# reproject the map 
sea2 <- spTransform( sea, CRS("+init=epsg:3395") )

# convert the sp map format to 
# an sf (simple features) format:
# ggmap requires the sf format
sea.sf <- st_as_sf( sea2 )

# separate out the data frame from the map
d <- as.data.frame( sea.sf )


```


Community Demographics  
=====================================  

Inputs {.sidebar}
-------------------------------------

```{r}
these.variables <- c("pnhwht12", "pnhblk12", "phisp12", "pntv12", "pfb12", "polang12", 
"phs12", "pcol12", "punemp12", "pflabf12", "pprof12", "pmanuf12", 
"pvet12", "psemp12", "hinc12", "incpc12", "ppov12", "pown12", 
"pvac12", "pmulti12", "mrent12", "mhmval12", "p30old12", "p10yrs12", 
"p18und12", "p60up12", "p75up12", "pmar12", "pwds12", "pfhh12")


# replace these with descriptive labels 
# from the data dictionary 
variable_labels <- c(
  "pnhwht12" = "Non-Hispanic White Population",
  "pnhblk12" = "Non-Hispanic Black Population",
  "phisp12" = "Hispanic Population",
  "pntv12" = "Native American Population",
  "pfb12" = "Foreign-Born Population",
  "polang12" = "Speaks Other Languages",
  "phs12" = "High School Education",
  "pcol12" = "College Education",
  "punemp12" = "Unemployment Rate",
  "pflabf12" = "Labor Force Participation",
  "pprof12" = "Professional Jobs",
  "pmanuf12" = "Manufacturing Jobs",
  "pvet12" = "Veteran Population",
  "psemp12" = "Self-Employment Rate",
  "hinc12" = "Household Income",
  "incpc12" = "Income Per Capita",
  "ppov12" = "Poverty Rate",
  "pown12" = "Home Ownership Rate",
  "pvac12" = "Vacancy Rate",
  "pmulti12" = "Multi-Unit Housing",
  "mrent12" = "Median Rent",
  "mhmval12" = "Median Home Value",
  "p30old12" = "Built Before 1930",
  "p10yrs12" = "Moved in Last 10 Years",
  "p18und12" = "Population Under 18",
  "p60up12" = "Population 60 and Older",
  "p75up12" = "Population 75 and Older",
  "pmar12" = "Married Population",
  "pwds12" = "With Disability",
  "pfhh12" = "Female-Headed Households"
)
temp.names <- paste0( "2012 ",variable_labels)

radioButtons( inputId="demographics", 
              label = h3("Census Variables"),
              # choices = variable_labels, 
              choiceNames= temp.names,
              choiceValues= these.variables,
              selected="pnhwht12")


# Adding interpretable variable names
# from the data dictionary:
# add a name attribute for each variable
# 
# value <- c(1,2,3)
# dd.name <- c("one","two","three")
# 
# x <- dd.name
# names(x) <- value
#
# dd names and values linked
# names( x[2] )
#
# can now get the label using the value
# using the name attributes 
# x[ "two" ]
#
# to add labels to the maps
# use the radio button value 
# to get the data dictionary label: 
#
# x[ input$demographics ]
```



Row {.tabset}
-------------------------------------


### Choropleth Map


```{r}

renderPlot({

# split the selected variable into deciles 
  
get_data <- 
  reactive({
             sea.sf <- 
             sea.sf %>% 
             mutate( q = ntile( get(input$demographics), 10 ) )  
          })

  
ggplot( get_data() ) +
    geom_sf( aes( fill = q ), color=NA ) +
    coord_sf( datum=NA ) +
    labs( title = paste0( "Choropleth of Select Demographics: ", toupper(input$demographics) ),
          caption = "Source: Harmonized Census Files",
          fill = "Population Deciles" ) +
    scale_fill_gradientn( colours=rev(ocean.balance(10)), guide = "colourbar" ) + 
    xlim( xmin = -13700000, xmax = -13500000 ) + 
    ylim( ymin = 5900000, ymax = 6100000 )

})

```


### Variable Distribution 

```{r}
renderPlot({

# extract vector x from the data frame 
# x <-  d[ "pnhwht12" ] %>% unlist()

  
get_variable_x <- reactive({ d[ input$demographics ] })

x <- get_variable_x() %>% unlist()

cut.points <- quantile( x, seq( 0, 1, 0.1 ) )

hist( x, breaks=50, 
      col="gray", border="white", yaxt="n",
      main=paste0( "Histogram of variable ", toupper( input$demographics ) ),
      xlab="red lines represent decile cut points" )

abline( v=cut.points, col="darkred", lty=3, lwd=2 )


})
```




Neighborhoods  
===================================== 

### Clusters  

```{r}

# define the bounding box corners 
bb <- st_bbox( c( xmin = -13700000, xmax = -13500000, 
                  ymax = 6100000, ymin = 5900000 ), 
               crs = st_crs("+init=epsg:3395"))


# ADD YOUR CUSTOM LABELS TO THE CLUSTERS 

sea2$cluster[ sea2$cluster == "1" ] <- "Veteran Natives in Vacant Homes"
sea2$cluster[ sea2$cluster == "2" ] <- "Women-led Multilingual Households"
sea2$cluster[ sea2$cluster == "3" ] <- "Established Professionals with Degrees"
sea2$cluster[ sea2$cluster == "4" ] <- "Ethnically Diverse Blue Collar Workers"


renderTmap({

tmap_mode("view")
tm_basemap( "CartoDB.Positron" )
tm_shape( sea2, bbox=bb ) + 
  tm_polygons( col="cluster", palette="-viridis",
               title="Community Types" ) 

})
```




NH Change 2000-2010  
===================================== 



Inputs {.sidebar}
-------------------------------------

```{r}

button.labels <- c("Median Home Value 2000","Median Home Value 2010","Value Change 2000-2010","Growth in Home Value")
button.values <- c("mhv.00","mhv.12","mhv.change","mhv.growth")
  
radioButtons( inputId="home.value", 
              label = h3("Home Values"),
              # choices = these.variables, 
              choiceNames=button.labels,
              choiceValues=button.values,
              selected="mhv.00")

```




Row {.tabset}
-------------------------------------



### Median Home Values


```{r}

renderPlot({

# split the selected variable into deciles 
  
get_data <- 
  reactive({
             sea.sf <- 
             sea.sf %>% 
             mutate( q = ntile( get(input$home.value), 10 ) )  
          })

  
ggplot( get_data() ) +
    geom_sf( aes( fill = q ), color=NA ) +
    coord_sf( datum=NA ) +
    labs( title = paste0( "Spatial Distribution of Home Values: ", toupper(input$demographics) ),
          caption = "Source: Harmonized Census Files",
          fill = "Home Value Deciles" ) +
    scale_fill_gradientn( colours=rev(ocean.balance(10)), guide = "colourbar" ) + 
    xlim( xmin = -13700000, xmax = -13500000 ) + 
    ylim( ymin = 5900000, ymax = 6100000 )

})

```


### Variable Distribution 

```{r}
renderPlot({

# extract vector x from the data frame 
# x <-  d[ "pnhwht12" ] %>% unlist()

  
get_variable_x <- reactive({ d[ input$home.value ] })

x <- get_variable_x() %>% unlist() %>% as.numeric()

cut.points <- quantile( x, seq( 0, 1, 0.1 ) )

hist( x, breaks=50, 
      col="gray", border="white", yaxt="n",
      main=paste0( "Histogram of ", toupper( input$home.value ) ),
      xlab="red lines represent decile cut points" )

abline( v=cut.points, col="darkred", lty=3, lwd=2 )


})
```  







Drivers of Change   
===================================== 



Inputs {.sidebar}
-------------------------------------

```{r}

button.labels <- c("Median Home Value 2000","Median Home Value 2010","Value Change 2000-2010","Growth in Home Value")
button.values <- c("mhv.00","mhv.12","mhv.change","mhv.growth")
  
radioButtons( inputId="dv", 
              label = h3("Select Your Dependent Variable"),
              choiceNames=button.labels,
              choiceValues=button.values,
              selected="mhv.change")


covariates <- c("pnhwht12", "pnhblk12", "phisp12", "pntv12", "pfb12", "polang12", 
"phs12", "pcol12", "punemp12", "pflabf12", "pprof12", "pmanuf12", 
"pvet12", "psemp12", "hinc12", "incpc12", "ppov12", "pown12", 
"pvac12", "pmulti12", "mrent12", "mhmval12", "p30old12", "p10yrs12", 
"p18und12", "p60up12", "p75up12", "pmar12", "pwds12", "pfhh12")

# covariate.labels <- c( ... )

checkboxGroupInput( inputId="covariates", 
              label = h3("Select Variables for Your Model"),
              choices = covariates,
              # choiceNames=covariate.labels,
              # choiceValues=covariates,
              selected=c("pnhwht12","pprof12","pvac12") )

```




Row {.tabset}
-------------------------------------



### Predicting Change 

```{r, results="asis"}

# RUNNING A REGRESSION WITH USER INPUTS
#
# create a formula object
# by constructing the formula from user selections 
# as a string then casting as a formula object 
#x.s <- c("x1","x2","x3" )
#formula.text <- paste( "y", " ~ ", paste0( x.s, collapse=" + ") )
#formula.object <- as.formula( formula.text )
#
#lm( formula.object, data=sea2 )
#
# make sure all variables are in d


  
# check boxes return vectors 
get_covariates <-  reactive({ input$covariates })


renderUI({
  
covariates <- get_covariates() 

formula.text <- paste0( "mhmval12", " ~ ", paste( covariates, collapse=" + " ) )
fo <- as.formula( formula.text )

m <- lm( fo, data=d )


# HTML( "<div style='width: 60%; margin: 0px auto;'>" )
HTML( 
  
  c("<br><br><br>",
  "<div type='regression' style='width: 60%; margin: 0px auto;'>",
  stargazer( m, type="html", omit.stat=c("rsq","f") ),
  "</div>",
  "<br><br><br>"
  )
  
  )


})

# HTML( reg.table )
```



### Correlation Plots 

```{r}
pairs( iris )
```





<style>

.chart-shim { overflow: auto; }
 
table{
   border-spacing:1px;
   margin-top:30px;
   margin-bottom:30px;
   margin-left: auto;
   margin-right: auto;
   align:center} 

td{ padding: 6px 10px 6px 10px } 

th{ text-align: left; } 

</style>